---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SwanScorer

<!-- badges: start -->

<!-- badges: end -->

The Strengths and Weaknesses of ADHD Symptoms and Normal Behavior Rating
Scale (SWAN) is a validated instrument for measuring
attention-deficit/hyperactivity disorder (ADHD) traits ([Burton et al.,
2018](https://doi.org/10.1101/248484)). The SwanScorer package provides
an easy way to automatically score your SWAN tests.

## Instructions

### Formatting Your Data

Our first step is to prepare your raw SWAN data.

1.  Prepare a spreadsheet, preferably a .csv file, with a row for each
    of the tests you'd like to score.

2.  Be sure the following columns are present in the spreadsheet and
    rename the columns to match the reference guide below. All of the
    following columns are necessary for the model. If age, gender, or
    parent_respondent are missing from a row, the model will not return
    a t-score for that row.

    1.  <div>

        +-----------------+-------------------+-------------------+
        | Column Name     | Additional        | Values            |
        |                 | Information       |                   |
        +=================+===================+===================+
        | age             | The child's age   | 5 - 18            |
        +-----------------+-------------------+-------------------+
        | gender          | The child's       | 1 = Male          |
        |                 | gender            |                   |
        |                 |                   | 2 = Female        |
        +-----------------+-------------------+-------------------+
        | pa              | Whether the       | 1 = Parent        |
        | rent_respondent | survey was filled | Respondent        |
        |                 | out by the parent |                   |
        |                 | or the child      | 0 = Child / Youth |
        |                 |                   | Self Respondent   |
        +-----------------+-------------------+-------------------+
        | swan1           | 1\. Give close    | -3 (Far Below) to |
        |                 | attention to      | 3 (Far Above)     |
        |                 | detail and avoid  |                   |
        |                 | careless mistakes |                   |
        +-----------------+-------------------+-------------------+
        | swan2           | 2\. Sustain       | -3 (Far Below) to |
        |                 | attention on      | 3 (Far Above)     |
        |                 | tasks and play    |                   |
        |                 | activities        |                   |
        +-----------------+-------------------+-------------------+
        | swan3           | 3\. Listen when   | -3 (Far Below) to |
        |                 | spoken to         | 3 (Far Above)     |
        |                 | directly          |                   |
        +-----------------+-------------------+-------------------+
        | swan4           | 4\. Follow        | -3 (Far Below) to |
        |                 | through on        | 3 (Far Above)     |
        |                 | instructions and  |                   |
        |                 | finish schoolwork |                   |
        |                 | / chores          |                   |
        +-----------------+-------------------+-------------------+
        | swan5           | 5\. Organize      | -3 (Far Below) to |
        |                 | tasks and         | 3 (Far Above)     |
        |                 | activities        |                   |
        +-----------------+-------------------+-------------------+
        | swan6           | 6\. Engage in     | -3 (Far Below) to |
        |                 | tasks that        | 3 (Far Above)     |
        |                 | require sustained |                   |
        |                 | mental effort     |                   |
        +-----------------+-------------------+-------------------+
        | swan7           | 7\. Keep track of | -3 (Far Below) to |
        |                 | things necessary  | 3 (Far Above)     |
        |                 | for activities    |                   |
        +-----------------+-------------------+-------------------+
        | swan8           | 8\. Ignore        | -3 (Far Below) to |
        |                 | extraneous        | 3 (Far Above)     |
        |                 | stimuli (able to  |                   |
        |                 | ignore background |                   |
        |                 | n                 |                   |
        |                 | o                 |                   |
        |                 | ise/distractions) |                   |
        +-----------------+-------------------+-------------------+
        | swan9           | 9\. Remember      | -3 (Far Below) to |
        |                 | daily activities  | 3 (Far Above)     |
        +-----------------+-------------------+-------------------+
        | swan10          | 10\. Sit still    | -3 (Far Below) to |
        |                 | (control          | 3 (Far Above)     |
        |                 | movements of      |                   |
        |                 | hands / feet or   |                   |
        |                 | control           |                   |
        |                 | squirming)        |                   |
        +-----------------+-------------------+-------------------+
        | swan11          | 11\. Stay seated  | -3 (Far Below) to |
        |                 | (when required by | 3 (Far Above)     |
        |                 | class rules /     |                   |
        |                 | social            |                   |
        |                 | conventions)      |                   |
        +-----------------+-------------------+-------------------+
        | swan12          | 12\. Modulate     | -3 (Far Below) to |
        |                 | motor activity    | 3 (Far Above)     |
        |                 | (inhibit          |                   |
        |                 | inappropriate     |                   |
        |                 | running /         |                   |
        |                 | climbing)         |                   |
        +-----------------+-------------------+-------------------+
        | swan13          | 13\. Play quietly | -3 (Far Below) to |
        |                 | (keep noise level | 3 (Far Above)     |
        |                 | reasonable)       |                   |
        +-----------------+-------------------+-------------------+
        | swan14          | 14\. Settle down  | -3 (Far Below) to |
        |                 | and rest (control | 3 (Far Above)     |
        |                 | constant          |                   |
        |                 | activity)         |                   |
        +-----------------+-------------------+-------------------+
        | swan15          | 15\. Modulate     | -3 (Far Below) to |
        |                 | verbal activity   | 3 (Far Above)     |
        |                 | (control excess   |                   |
        |                 | talking)          |                   |
        +-----------------+-------------------+-------------------+
        | swan16          | 16\. Reflect on   | -3 (Far Below) to |
        |                 | questions         | 3 (Far Above)     |
        |                 | (control blurting |                   |
        |                 | out answers)      |                   |
        +-----------------+-------------------+-------------------+
        | swan17          | 17\. Await turn   | -3 (Far Below) to |
        |                 | (stand in line    | 3 (Far Above)     |
        |                 | and take turns)   |                   |
        +-----------------+-------------------+-------------------+
        | swan18          | 18\. Enter into   | -3 (Far Below) to |
        |                 | conversations and | 3 (Far Above)     |
        |                 | games (control    |                   |
        |                 | interrupting /    |                   |
        |                 | intruding)        |                   |
        +-----------------+-------------------+-------------------+

        </div>

3.  Any additional columns in the spreadsheet, i.e. an identifier, will
    remain untouched in the spreadsheet

### Installation

You can install the development version of SwanScorer from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jclutton/SwanScorer")
```

### Generate Scores

Use the code below to generate your t-scores. First, You will be
prompted to select your file. Second, we will check that your data are
formatted properly. And third, t-scores for the full test as well as the
two subdomains (inattentive and hyperactive) will all be generated

```{r example, eval = FALSE}
library(SwanScorer)

# The get_swan_tscores checks that your data are formatted correctly and generates the t-scores
swan_tscores <- get_swan_tscores()
```

### Additional options

You have the option to specify...

1.  the file path of the input file

2.  a folder to save a output spreadsheet with the t-scores

```{r example2, eval=FALSE}
library(here)
library(SwanScorer)

# Example of how to specify the input file
swan_tscores <- get_swan_tscores(file = here("test_scores.csv"))

# Example of how to automatically save a spreadsheet
swan_tscores <- get_swan_tscores(output_folder = here())


```

## Understanding the Output
