---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SwanScorer

<!-- badges: start -->

<!-- badges: end -->

The Strengths and Weaknesses of ADHD Symptoms and Normal Behavior Rating
Scale (SWAN) is a validated instrument for measuring
attention-deficit/hyperactivity disorder (ADHD) traits ([Burton et al.,
2018](https://doi.org/10.1101/248484)). The SwanScorer package provides
an easy way to automatically score your SWAN tests.

## Instructions

### Formatting Your Data

Our first step is to prepare your raw SWAN data.

1.  Prepare a spreadsheet, preferably a .csv file, with a row for each
    of the tests you'd like to score.

2.  Be sure the following columns are present in the spreadsheet and
    rename the columns to match the reference guide below. All of the
    following columns are necessary for the model. If age, gender, or
    p_respondent are missing from a row, the model will not return a
    t-score for that row.

    <div>

    +--------------+--------------+--------------+
    | Column Name  | Additional   | Values       |
    |              | Information  |              |
    +==============+==============+==============+
    | age          | The child's  | 5 - 18       |
    |              | age          |              |
    +--------------+--------------+--------------+
    | gender       | The child's  | 1 = Male     |
    |              | gender       |              |
    |              |              | 2 = Female   |
    +--------------+--------------+--------------+
    | p_respondent | Whether the  | 1 = Parent   |
    |              | survey was   | Respondent   |
    |              | filled out   |              |
    |              | by the       | 0 = Child /  |
    |              | parent or    | Youth Self   |
    |              | the child    | Respondent   |
    +--------------+--------------+--------------+
    | swan1        | 1\. Give     | -3 (Far      |
    |              | close        | Below) to 3  |
    |              | attention to | (Far Above)  |
    |              | detail and   |              |
    |              | avoid        |              |
    |              | careless     |              |
    |              | mistakes     |              |
    +--------------+--------------+--------------+
    | swan2        | 2\. Sustain  | -3 (Far      |
    |              | attention on | Below) to 3  |
    |              | tasks and    | (Far Above)  |
    |              | play         |              |
    |              | activities   |              |
    +--------------+--------------+--------------+
    | swan3        | 3\. Listen   | -3 (Far      |
    |              | when spoken  | Below) to 3  |
    |              | to directly  | (Far Above)  |
    +--------------+--------------+--------------+
    | swan4        | 4\. Follow   | -3 (Far      |
    |              | through on   | Below) to 3  |
    |              | instructions | (Far Above)  |
    |              | and finish   |              |
    |              | schoolwork / |              |
    |              | chores       |              |
    +--------------+--------------+--------------+
    | swan5        | 5\. Organize | -3 (Far      |
    |              | tasks and    | Below) to 3  |
    |              | activities   | (Far Above)  |
    +--------------+--------------+--------------+
    | swan6        | 6\. Engage   | -3 (Far      |
    |              | in tasks     | Below) to 3  |
    |              | that require | (Far Above)  |
    |              | sustained    |              |
    |              | mental       |              |
    |              | effort       |              |
    +--------------+--------------+--------------+
    | swan7        | 7\. Keep     | -3 (Far      |
    |              | track of     | Below) to 3  |
    |              | things       | (Far Above)  |
    |              | necessary    |              |
    |              | for          |              |
    |              | activities   |              |
    +--------------+--------------+--------------+
    | swan8        | 8\. Ignore   | -3 (Far      |
    |              | extraneous   | Below) to 3  |
    |              | stimuli      | (Far Above)  |
    |              | (able to     |              |
    |              | ignore       |              |
    |              | background n |              |
    |              | o i s e /    |              |
    |              | d            |              |
    |              | istractions) |              |
    +--------------+--------------+--------------+
    | swan9        | 9\. Remember | -3 (Far      |
    |              | daily        | Below) to 3  |
    |              | activities   | (Far Above)  |
    +--------------+--------------+--------------+
    | swan10       | 10\. Sit     | -3 (Far      |
    |              | still        | Below) to 3  |
    |              | (control     | (Far Above)  |
    |              | movements of |              |
    |              | hands / feet |              |
    |              | or control   |              |
    |              | squirming)   |              |
    +--------------+--------------+--------------+
    | swan11       | 11\. Stay    | -3 (Far      |
    |              | seated (when | Below) to 3  |
    |              | required by  | (Far Above)  |
    |              | class rules  |              |
    |              | / social     |              |
    |              | conventions) |              |
    +--------------+--------------+--------------+
    | swan12       | 12\.         | -3 (Far      |
    |              | Modulate     | Below) to 3  |
    |              | motor        | (Far Above)  |
    |              | activity     |              |
    |              | (inhibit     |              |
    |              | i            |              |
    |              | nappropriate |              |
    |              | running /    |              |
    |              | climbing)    |              |
    +--------------+--------------+--------------+
    | swan13       | 13\. Play    | -3 (Far      |
    |              | quietly      | Below) to 3  |
    |              | (keep noise  | (Far Above)  |
    |              | level        |              |
    |              | reasonable)  |              |
    +--------------+--------------+--------------+
    | swan14       | 14\. Settle  | -3 (Far      |
    |              | down and     | Below) to 3  |
    |              | rest         | (Far Above)  |
    |              | (control     |              |
    |              | constant     |              |
    |              | activity)    |              |
    +--------------+--------------+--------------+
    | swan15       | 15\.         | -3 (Far      |
    |              | Modulate     | Below) to 3  |
    |              | verbal       | (Far Above)  |
    |              | activity     |              |
    |              | (control     |              |
    |              | excess       |              |
    |              | talking)     |              |
    +--------------+--------------+--------------+
    | swan16       | 16\. Reflect | -3 (Far      |
    |              | on questions | Below) to 3  |
    |              | (control     | (Far Above)  |
    |              | blurting out |              |
    |              | answers)     |              |
    +--------------+--------------+--------------+
    | swan17       | 17\. Await   | -3 (Far      |
    |              | turn (stand  | Below) to 3  |
    |              | in line and  | (Far Above)  |
    |              | take turns)  |              |
    +--------------+--------------+--------------+
    | swan18       | 18\. Enter   | -3 (Far      |
    |              | into         | Below) to 3  |
    |              | c            | (Far Above)  |
    |              | onversations |              |
    |              | and games    |              |
    |              | (control     |              |
    |              | interrupting |              |
    |              | / intruding) |              |
    +--------------+--------------+--------------+

    </div>

3.  Any additional columns in the spreadsheet, i.e. an identifier, will
    remain untouched in the spreadsheet

### Installation

You can install the development version of SwanScorer from
[GitHub](https://github.com/jclutton/SwanScorer) with:

```{r install, eval = FALSE }
# If the pak package isn't installed already, do so first. 
install.packages("pak")

pak::pak("jclutton/SwanScorer")
```

### Generate Scores

Use the code below to generate your t-scores. First, You will be
prompted to select your file. Second, we will check that your data are
formatted properly. And third, t-scores for the full test as well as the
two subdomains (inattentive and hyperactive) will all be generated

```{r example, eval = FALSE}
library(SwanScorer)

# The get_swan_tscores checks that your data are formatted correctly and generates the t-scores
swan_tscores <- get_swan_tscores()
```

### Additional options

You have the option to specify...

1.  the file path of the input file

2.  a folder to save a output spreadsheet with the t-scores

```{r example2, eval=FALSE}
library(here)
library(SwanScorer)

# Example of how to specify the input file
swan_tscores <- get_swan_tscores(file = here("test_scores.csv"))

# Example of how to automatically save a spreadsheet
swan_tscores <- get_swan_tscores(output_folder = here())


```

## Understanding the Output

+----------------------------------+----------------------------------+
| Column                           | Description                      |
+==================================+==================================+
| swan_gender_tscores              | A t-score across the full SWAN   |
|                                  | test that adjusts for age,       |
|                                  | respondent, and gender           |
+----------------------------------+----------------------------------+
| swan_tscores                     | A t-score across the full SWAN   |
|                                  | test that adjusts for age and    |
|                                  | respondent                       |
+----------------------------------+----------------------------------+
| swan_ia_gender_tscores           | A t-score of the inattentive     |
|                                  | subdomain (questions 1-9) that   |
|                                  | adjusts for age, respondent, and |
|                                  | gender                           |
+----------------------------------+----------------------------------+
| swan_ia_tscores                  | A t-score of the inattentive     |
|                                  | subdomain (questions 1-9) that   |
|                                  | adjusts for age and respondent   |
+----------------------------------+----------------------------------+
| swan_hi_gender_tscores           | A t-score of the hyperactive     |
|                                  | subdomain (questions 10-18) that |
|                                  | adjusts for age, respondent, and |
|                                  | gender                           |
+----------------------------------+----------------------------------+
| swan_hi_tscores                  | A t-score of the hyperactive     |
|                                  | subdomain (questions 10-18) that |
|                                  | adjusts for age and respondent   |
+----------------------------------+----------------------------------+
