---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SwanScorer

<!-- badges: start -->

<!-- badges: end -->

The Strengths and Weaknesses of ADHD Symptoms and Normal Behavior Rating
Scale (SWAN) is a validated instrument for measuring
attention-deficit/hyperactivity disorder (ADHD) traits ([Burton et al.,
2018](https://doi.org/10.1101/248484)). The SwanScorer package provides
an easy way to automatically score your SWAN tests.

## Instructions

### Formatting Your Data

Our first step is to prepare your raw SWAN data.

1.  Prepare a spreadsheet, preferably a .csv file, with a row for each
    of the tests you'd like to score.

2.  Be sure the following columns are present in the spreadsheet and
    rename the columns to match the reference guide below. All of the
    following columns are necessary for the model. If age, gender, or
    parent_respondent are missing from a row, the model will not return
    a t-score for that row.

    1.  <div>

        +------------------+----------------------+----------------------+
        | Column Name      | Additional           | Values               |
        |                  | Information          |                      |
        +==================+======================+======================+
        | age              | The child's age      | 5 - 18               |
        +------------------+----------------------+----------------------+
        | gender           | The child's gender   | 1 = Male             |
        |                  |                      |                      |
        |                  |                      | 2 = Female           |
        +------------------+----------------------+----------------------+
        | p                | Whether the survey   | 1 = Parent           |
        | arent_respondent | was filled out by    | Respondent           |
        |                  | the parent or the    |                      |
        |                  | child                | 0 = Child / Youth    |
        |                  |                      | Self Respondent      |
        +------------------+----------------------+----------------------+
        | swan1            | 1\. Give close       | -3 (Far Below) to 3  |
        |                  | attention to detail  | (Far Above)          |
        |                  | and avoid careless   |                      |
        |                  | mistakes             |                      |
        +------------------+----------------------+----------------------+
        | swan2            | 2\. Sustain          | -3 (Far Below) to 3  |
        |                  | attention on tasks   | (Far Above)          |
        |                  | and play activities  |                      |
        +------------------+----------------------+----------------------+
        | swan3            | 3\. Listen when      | -3 (Far Below) to 3  |
        |                  | spoken to directly   | (Far Above)          |
        +------------------+----------------------+----------------------+
        | swan4            | 4\. Follow through   | -3 (Far Below) to 3  |
        |                  | on instructions and  | (Far Above)          |
        |                  | finish schoolwork /  |                      |
        |                  | chores               |                      |
        +------------------+----------------------+----------------------+
        | swan5            | 5\. Organize tasks   | -3 (Far Below) to 3  |
        |                  | and activities       | (Far Above)          |
        +------------------+----------------------+----------------------+
        | swan6            | 6\. Engage in tasks  | -3 (Far Below) to 3  |
        |                  | that require         | (Far Above)          |
        |                  | sustained mental     |                      |
        |                  | effort               |                      |
        +------------------+----------------------+----------------------+
        | swan7            | 7\. Keep track of    | -3 (Far Below) to 3  |
        |                  | things necessary for | (Far Above)          |
        |                  | activities           |                      |
        +------------------+----------------------+----------------------+
        | swan8            | 8\. Ignore           | -3 (Far Below) to 3  |
        |                  | extraneous stimuli   | (Far Above)          |
        |                  | (able to ignore      |                      |
        |                  | background           |                      |
        |                  | noise/distractions)  |                      |
        +------------------+----------------------+----------------------+
        | swan9            | 9\. Remember daily   | -3 (Far Below) to 3  |
        |                  | activities           | (Far Above)          |
        +------------------+----------------------+----------------------+
        | swan10           | 10\. Sit still       | -3 (Far Below) to 3  |
        |                  | (control movements   | (Far Above)          |
        |                  | of hands / feet or   |                      |
        |                  | control squirming)   |                      |
        +------------------+----------------------+----------------------+
        | swan11           | 11\. Stay seated     | -3 (Far Below) to 3  |
        |                  | (when required by    | (Far Above)          |
        |                  | class rules / social |                      |
        |                  | conventions)         |                      |
        +------------------+----------------------+----------------------+
        | swan12           | 12\. Modulate motor  | -3 (Far Below) to 3  |
        |                  | activity (inhibit    | (Far Above)          |
        |                  | inappropriate        |                      |
        |                  | running / climbing)  |                      |
        +------------------+----------------------+----------------------+
        | swan13           | 13\. Play quietly    | -3 (Far Below) to 3  |
        |                  | (keep noise level    | (Far Above)          |
        |                  | reasonable)          |                      |
        +------------------+----------------------+----------------------+
        | swan14           | 14\. Settle down and | -3 (Far Below) to 3  |
        |                  | rest (control        | (Far Above)          |
        |                  | constant activity)   |                      |
        +------------------+----------------------+----------------------+
        | swan15           | 15\. Modulate verbal | -3 (Far Below) to 3  |
        |                  | activity (control    | (Far Above)          |
        |                  | excess talking)      |                      |
        +------------------+----------------------+----------------------+
        | swan16           | 16\. Reflect on      | -3 (Far Below) to 3  |
        |                  | questions (control   | (Far Above)          |
        |                  | blurting out         |                      |
        |                  | answers)             |                      |
        +------------------+----------------------+----------------------+
        | swan17           | 17\. Await turn      | -3 (Far Below) to 3  |
        |                  | (stand in line and   | (Far Above)          |
        |                  | take turns)          |                      |
        +------------------+----------------------+----------------------+
        | swan18           | 18\. Enter into      | -3 (Far Below) to 3  |
        |                  | conversations and    | (Far Above)          |
        |                  | games (control       |                      |
        |                  | interrupting /       |                      |
        |                  | intruding)           |                      |
        +------------------+----------------------+----------------------+

        </div>

3.  Any additional columns in the spreadsheet, i.e. an identifier, will
    remain untouched in the spreadsheet

### Installation

You can install the development version of SwanScorer from
[GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jclutton/SwanScorer")
```

### Generate Scores

Use the code below to generate your t-scores. First, You will be
prompted to select your file. Second, we will check that your data are
formatted properly. And third, t-scores for the full test as well as the
two subdomains (inattentive and hyperactive) will all be generated

```{r example, eval = FALSE}
library(SwanScorer)

# The get_swan_tscores checks that your data are formatted correctly and generates the t-scores
swan_tscores <- get_swan_tscores()
```

### Additional options

You have the option to specify...

1.  the file path of the input file

2.  a folder to save a output spreadsheet with the t-scores

```{r example2, eval=FALSE}

# Example of how to specify the input file
swan_tscores <- get_swan_tscores(file = here("test_scores.csv"))

# Example of how to automatically save a spreadsheet
swan_tscores <- get_swan_tscores(output_folder = here())


```

